---
title: "Proposal"
author: "Brett Chrisler"
date: "April 9, 2018"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
bibliography: Bibliography.xml
---

<!--***DAN: Search ***DAN in all docs to find my comments-->

<!--***DAN: The proposal: There are various distracting typos and gramatical errors. The proposal is basically fine to get you started, but it is pretty non-explicit in what you will actually do/calculate, and it is also somewhat limited in what you seem to directly have in mind. You need to specify some secific research questions and hypotheses. Just counting MJO events and major events is a minimal project and I suggets you formulate a concrete plan for how to go beyond that. We spoke about spectral analyses but none of that is in there. Whereas if you just do what you have proposed, you will probably do fine, grade-wise, you can do much better than that and it will help you learn if you do. The place to start is scientific questions.-->

<!--***DAN: Overall, this meets the criteria of an effort-full attempt at all parts of the assignment. 10/10.-->


```{r find_fixed_source, echo=F}
source("find_active.R")
source("find_ocur.R")
source("find_case.R")
```

<!--***DAN: When I tried to knit, it did not work, the first error I got was missing packages which were not installed on my machine. Rather than go through and install packages, I put the below chunk in, which will find and install the 2018-03-29 version (you can change that date) of all packages you use in a local folder (called .checkpoint, I told the machine to put it in the repo root directory, but of course it will not be tracked by git unless you add it, and you should not). I have found that to be a good place to put it because that is the one directory location that you know will exist on other people's machines. For some reason it still did not work, and it indicted the problem was with the xtable package. Apparently it was failing to find that you had made a call to library(xtable) below. When I added library(xtable) here it worked and knitted. This checkpoint business can improve portability of your code by installing packages for you on other people's machnes, in addition to making the analysis reproducible through time even if the packages are changed. I hope this helps.  
<!--checkpoint package-->
```{r checkpoint_chunk, echo=F, warning=F, message=F, results="hide"}
library(checkpoint)
library(xtable)
if (!dir.exists("./.checkpoint/")){
  dir.create("./.checkpoint/")
}
checkpoint("2018-03-29",checkpointLocation = "./")
```

```{r test active, echo=F}
#install.packages("xtable")
data <-read.csv("RMM.csv")
ind <- as.numeric(data$amplitude)
min = 1
ocur <- 30
h <- find_active(ind,min)
case <- as.matrix(find_ocur(h))
final <- as.matrix(find_case(case,ocur))

```

# __Comparing Indices: Madden-Julian Oscillation and Termination__

### I. Motivation
The Madden - Julian Oscillation (MJO) is a large-scale circulation that is occasionally coupled with convection. The MJO propagates eastward at around 5 m/s and has an intraseasonal lifespan of 30 - 90 days. The MJO most often begins in the Indian Ocean and ends in the Western Pacific; however, location can vary immensely and is divided into eight phases. There are two modes of the MJO which are active and inactive. During the active mode, strong convection prevails and higher precipitation amounts are observed, while the inactive mode. To analyze the length of an active mode, indices are created to set a threshold of when the MJO is either mode [@Zhang2005].

Comparing these indices to see how they do well with recognizing the active mode, as well as when an event is terminated is crucial for the study of the Madden-Julian Oscillation. The birth and decay of the MJO is still unknown and comparing indices could shine light on what precursor conditions allow for the MJO to maintain itself. As the MJO has widespread impacts on global climate, understanding this phenomena is important especially in regards to global climate models (GCMs).
<br>

### II. Data and Analysis
Multiple indcies exist for the Madden-Julian Oscillation with the most used being the Real-time Multivariate MJO (RMM) index. This index uses both OLR and zonal winds to calcuate the principal components and amplitude of the MJO. Additional indices include the following: the OLR MJO Index, OMI, the Original OLR MJO Index, OOMI, the Real-time OLR MJO Index, RMP, the Filtered OLR MJO index, FMO, and the Velocity Potential MJO index, VPM. OMI, OOMI, RMO, and FMO are univariate and use OLR data, while the VPM is multivariate like the RMM but instead used 200 hPa velocity potentional instead of OLR. 
<br>

### III. Preliminary and Expected Results
Preliminary results that were calculated included the amount of events based off any days of amplitude greater than unity, the amount of days of amplitude above unity, and the amount of major events meaning maintaining longer than 30 days.

```{r xtable, echo=F}
library(xtable)
Category <-c("Amount of Active Modes", "Amount of Days of Active Modes", "Amount of Major Events (>30 days)")
Amount  <- c(length(case),sum(case),length(final))
tableinfo <- data.frame(Category, Amount)
xtable(tableinfo)
```


Results should at least be consistent among filters using similar methods, such as all OLR univariate indices should return similar results. Variations may occur due to different filtering among the indices. If multivariate results return similar results to the univariate ones, then it should indicate that the additional parameter is not necessarily important for maintaining an MJO event. Additionally, the length of an event will be a good indicator if a certain index is picking up on a known case study event.